---
title: "Analysis"
author: "Ashley Harris"
date: "4/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(MASS)
knitr::opts_chunk$set(echo = TRUE)
```

# Explanation of Linear Discriminant Analysis

# Exploring the data

```{r cars}
diabetes.df = read.csv("diabetes-bivar.csv", header=TRUE)
head(diabetes.df)
```

We have `dim(diabetes.df)[1]` participants in our dataset. 

The data are comprised of three columns:
*The "glutest" column represents the glucose tolerance test score for each participant.
*The "insulin" column represents the insulin present in the participant.
*The group column represents the classification for each participant: 1 for overt diabetes, 2 for chemical diabetes (no symptoms, but abnormal glucose tolerance test), and 3 for an overt diabetic.

## plotting the data

```{r}
plot(glutest~insulin, data=diabetes.df, col=group, main="Insulin levels vs glucose tolerance test scores", xlab="insulin level" ylab="glucose tolerance test score")
dim(diabetes.df)
```

# explanation of linear discriminant analysis (LDA)

LDA is a method for reducing dimensionality and highlighting differences between groups of observations.
*Dimensionality reduction aims to reduce the number of dimensions (i.e. variables) in a dataset while retaining as much information as possible.
*Highlighting differences between groups of observations allows us to predict an unknown characteristic of a new observation (in this case, whether a participant is diabetic or not) based on characteristics that we do know (insulin and glucose test results). 

# Conducting linear discriminant analysis

```{r}

diabetes.lda = lda(group~log(glutest)+insulin, data=diabetes)

```

